---
title: "Lab Report 1"
author: "Bingying Liu"
date: "January 13, 2019"
output: html_document
---

```{r setup, message=F, warning=F, echo=F}
library(tidyverse)
library(ggdark)
require(magrittr)
require(plyr)
ggplot2::theme_set(ggplot2::theme_bw())
knitr::opts_chunk$set(fig.align = 'center')
```

***

## Exercise 1

Create a code chunk and set the header parameter to TRUE and print out the top rows of the table with head() as above.


```{r}
data <- read.table(file = url("http://www2.stat.duke.edu/~pdh10/FCBS/Exercises/azdiabetes.dat"), header = TRUE)
head(data)
```


## Exercise 2

Generate a sequence of 100 equispaced real numbers from 0 to 1 and store it in a variable called seq2.

```{r}
seq2 <- seq(from = 0, to = 1, length.out = 101)
seq2
```

## Exercise 3

Sort the entries in seq3 from greatest to least.s

```{r}
seq3 <- seq(from = -3, to = 3, by = .5)
seq3 <- sort(seq3, decreasing = TRUE)
seq3
```

## Exercise 4

Find the variance of each row of mat5.

```{r}
# generate large matrix
mat5 <- matrix(seq(1,100,1), nrow = 4, ncol = 25, byrow = T)
apply(X= mat5, MARGIN = 1, FUN = var)

```

## Exercise 5

Generate 500 samples from a Beta distribution with shape parameter [a,b]=[0.5,0.5] and store the samples in a variable called W.

```{r}
w <- rbeta(500, shape1 = 0.5, shape2 = 0.5)
```


## Exercise 6

Browse online resources (some below), or use code from above to make a few plots of your own.

I used the mpg dataset (which is the fuel economy data from 1999 and 2008) to create some plots using ggplot darkmode.

```{r, message = F}
p <- ggplot(mpg) + 
  geom_point(aes(cty, hwy, color = class)) + 
  scale_y_continuous() +
  guides(color = guide_legend(reverse = TRUE)) +
  labs(title = "Highway fuel efficiency by City fuel efficiency and Type of car",
       x = "City fuel efficiency (miles per gallon)",
       y = "Highway fuel Efficiency (miles per gallon)",
       color = "Class")

p + dark_theme_gray() 
```


```{r}
p <- ggplot(mpg) + 
  geom_boxplot(aes(trans, hwy, color = trans)) + 
  coord_flip() +
  guides(color = guide_legend(reverse = TRUE)) +
  labs(title = "Highway fuel efficiency by type of transmission",
       x = "Type of transmission",
       y = "Highway fuel Efficiency (miles per gallon)")

p + dark_theme_gray() 
```
***

